### 2019-12-02 submit_jpb/reviewers/response.txt
include Z scaling work

Referee 1:

We thank the referee for the comments on our manuscript. The present work considers fully stripped ions. Indeed, this may be considered as a weakness of our calculations. Nonetheless, we plan to extend our theoretical approach to partially stripped ions in the future. Please find below our reply to his/her observations.

Detailed reply:

1) We have changed the wording of our findings in the corresponding sections.

2) The grammatical error was corrected.

3a) The dependence of the mean energy with the projectile charge at intermediate energy reflects the non-linear dependence on E and ZP of the continuum state in the CDW approximation. Then, the dependence of the numerator and the denominator on ZP is not canceled out, which explains the maximum in Fig. 2. On the contrary, the behavior at high energy follows the Z^2 rule, and the mean energy distribution tends to a universal value for all ions, as noted on page 3.
3b) We included a few sentences discussing the usefulness of understanding the relationship between the mean energy and angle distribution of secondary electrons with the radiation damage. 
3) We included a reference on radiation damage in the introduction as well.

4) The denominator in Eq. 4 is the same as in Eq. 3: the total cross-section. However, they are both integrated on different variables. We have included the corresponding expression of the normalization of the angular distribution.

5) We defined the dotted line accordingly.

6a) We included the reference to Table I in the caption.
6b) In general, the curves within each scaling band are ordered from the smallest to the largest. For the hydrocarbons, this behavior holds for both scalings since n_e for carbon does not change. However, this is not always the case (see, for example, adenine C5H5N5, uracil C4H4N2O2, and DNA backbone C5H10O5P in figure A attached). 
6c) Upon the suggestion of the referee, we included the Toburen numbers in Table II. The inclusion of the band by Toburen, together with the new scaling, was too confusing and finally decided not to modify the figure.

7) We included the Toburen scaling in figure B, attached. The x-axis refers to the number of active electrons, whether they are from the Toburen or the CDW scaling. As seen, the scaling by Toburen is not as good as ours, which follows Fig. 6. To avoid having too many curves, we decided not to modify Fig. 7, but we comment on these differences in the manuscript: 
"An equivalent plot using the Toburen numbers n_e of Table II does not exhibit the straight lines obtained with the present scaling."

8) We appreciate the suggestion of the Z-scaling at intermediate energies. We are currently investigating its implementation. We refer to it in the manuscript by including the sentences:
"At intermediate impact energies, the Z^2 rule no longer holds, and other scalings can be considered in this region. For example, the molecular cross section and ion impact energy can be reduced with the projectile charge Z, as suggested in [33, 34]."

Referee 2:

We decided to include in the introduction a comment about different previous work combining atomic cross sections to obtain molecular ones by Galassi et al. and Ludde et al.
"Different approaches have been proposed to deal with the ionization of molecular targets within the independent atom model. For example, Galassi et al. [2] obtain molecular cross sections by combining CDW-EIS atomic ones based on the population of the molecular orbitals. More recently, Lüdde et al. [3, 4] propose a combination of atomic cross sections with geometrical screening corrections."