### 2020-03-09 zscaling/main.tex
include zscaling paper

\documentclass[10pt,showpacs,showkeys,twocolumn]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \usepackage[T1]{fontspec}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{MnSymbol}
\usepackage{gensymb}
\usepackage{comment}
\usepackage{wasysym} 
\usepackage{stmaryrd}

\begin{document}

\title[Universal scaling for the ionization of biological molecules]{
Universal scaling for the ionization of biological molecules by highly charged ions}
\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET--UBA), \\ Buenos Aires, Argentina.}

\date{\today}% It is always \today, today,

\begin{abstract}
In the present work, we investigate scaling rules for the ionization cross sections of multicharged ions on molecules of biological interest. The cross sections are obtained from distorted--wave calculations for atomic targets
combined with stoichiometric model for the molecules proposed in [Mendez \textit{et al.} J. Phys B (2020)]. We examine ions with charges $Z$ from $+1$ to $+8$ in the pyrimide bases: adenine, citosine, guanine, thymine, uracil, tetrahydrofuran, pyrimidine, and also in water. We propose a scaling with the ion charge valid in the intermediate to high energy range, i.e. 0.2-5 MeV/amu for oxigen impact. We extend it to a universal scaling for any ion and molecule, which merges the forty ion--molecule systems analyzed here in a single band, and proved to be valid for other molecules too.  
\end{abstract}

\keywords{ionization, scaling, molecules, charged-ions, DNA, 
multicharged ions}
\pacs{34.50Gb, 34.80Gs, 34.80Dp}

\maketitle
%\ioptwocol

%\section{Introduction}
The ionization of biological molecules by multicharged ions has increasing interest due to medical and environmental reasons \cite{PhysMed}, from medical treatments  \cite{Mohamad2017,Solov2009,Denifl2011} to recognition of contaminants in biological materials \cite{water,ferrazdias}. Many semiempirical \citep{vera_prl2013} and  theoretical efforts are currently being done \cite{MendezJPB20,Quinto20,ludde2019,ludde2018,ludde2016,Champion2012} to get reliable values for the ionization cross sections of these molecules.

Recently we combined the continue distorted--wave calculations  (CDW) for atoms and the simple stoichiometric model (SSM) to approximate the ionization cross sections of complex molecular targets by charged ions \cite{MendezJPB20}. This CDW-SSM  approximation showed reasonable results for six ions in seventeen molecules. As expected, in the high energy range (i.e. above 5 MeV/amu) the ionization cross sections  have the $Z^2$ dependence predicted by the first Born approximation. Instead, at intermediate energies the dependence with $Z$ is more complex, and non-perturvative models are mandatory.

The aim of this letter is to deep on our previous work in \cite{MendezJPB20} by proposing a scaling of the ionization cross sections of complex molecules with the ion charge $Z$ valid at intermediate energies.  The requirements of values for a multiplicity of systems make this kind of scalings very interesting as a first order test for new experimental measurements and as inputs in multipurpose codes. Based on the scaling of the ionization cross sections with the number of active electrons, we propose an universal scaling for any ion-target system, valid at least for the complex biological molecules studied here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Scaling with the ion charge}
%\label{sec:Z-scaling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

At intermediate impact energies, the $Z^2$ rule no longer holds. Janev and Presnyakov ~\cite{janev1980} suggest $\sigma/Z$ versus $E/Z$  as the \textit{natural} reduced form of the ionization cross section $\sigma$. Much more recently, Montenegro and co-workers  ~\cite{dubois13,montenegro_pra13} proposed an alternative scaling  by taking into account that  $\sigma$ is a function of $Z^2/E$, so an scaling given by
\begin{equation}
    \sigma/Z^{\alpha}=f(E/Z^{2-\alpha}),
    \label{Montenegro}
\end{equation}
keeps the $Z^2/E$ relationship for any value of $\alpha$. 
%The scaling by Janev \cite{janev1980} displayed in figure {fig:zreduced}  is equivalent to using $\alpha=1$. 
In  \cite{dubois13} $\alpha=4/3$ is proposed for ionization of He and H$_2$ by different charged ions.

%\newpage
\begin{figure*}[!t]%[h]
\centering
\includegraphics[width=0.7\textwidth]{zscale_alpha.eps}
\caption{(Color online) Scaled ionization cross section $\sigma/Z^{\alpha}$ as a function 
of ion impact energy $E/Z^{2-\alpha}$ as explained in Eq. (\ref{Montenegro}). Colors are asociated with the impact charge as signed in the upper part of the figure. Curves: present CDW-SSM theoretical results.
Symbols: experimental impact of 
\mbox{\Large$\circ$} H$^+$~\cite{iriki2011} and 
$\lozenge$ C$^{+6}$ \cite{tribedi2019} on adenine;
H$^+$ on  $\triangle$ uracil~\cite{itoh2013}, 
$\ominus$ C$^{+4}$, 
$\oplus$ C$^{+6}$, O$^{+6}$, F$^{+6}$, and 
$\otimes$ O$^{+8}$, F$^{+8}$ on uracil~\cite{agnihotri2012,agnihotri2013};
H$^+$ on $\triangledown$ pyrimidine~\cite{wolff2014}, and
$\meddiamond$ THF~\cite{wang2016};
\mbox{\fontsize{10}{20} $\varhexagon$} H$^+$ \cite{Luna2007}, 
$\pentagon$ H$^+$ \cite{Rudd86}, 
$\odot$ He$^{+2}$ \cite{Rudd85},  
$\logof$ Li$^{+3}$ \cite{Luna_Li_water},  
$\ovee$ C$^{+6}$ \cite{DalCappello2009}, and 
$\varobslash$ O$^{+8}$ \cite{Tribedi_O_water} on water.
Markers~$\square$~\cite{rahman2016}, 
$\rhd$~\cite{bug2017}, 
$\lhd$~\cite{wolf2019}, and 
$\medstar$~\cite{fuss2009} correspond to electron impact ionization with equi-velocity conversion.}
\label{fig:zreduced}
\end{figure*}

% We consider our recent CDW-SSM results \cite{MendezJPB20} for eight targets (DNA and RNA nucleo-basis, tetrahydrofuran (THF) and pyrimidine) and five charged ions (H$^+$m He$^+2$, Be$^+4$, C$^+6$, and O$^+8$)eight targets (DNA and RNA nucleo-basis, tetrahydrofuran (THF) and pyrimidine) and five charged ions (H$^+$m He$^+2$, Be$^+4$, C$^+6$, and O$^+8$) as benchmark for the present scaling.

Combining Eq. (\ref{Montenegro}) and our recent CDW-SSM results \cite{MendezJPB20}, we propose here a $Z$-scaling and tested it %the best possible the cross sections of
for eight targets (the DNA and RNA nucleo-basis: adenine, cytosine, guanine, thymine, uracil, also tetrahydrofuran (THF), pyrimidine, and water) and five charged ions (H$^+$m He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$). We considered these forty collisional systems as benchmark for the present scaling.

%Based on these ion-target combinations, 
We found that $\sigma/Z^{\alpha}$ vs. $E^{2-\alpha}$, with $\alpha=1.2$ best fits the CDW-SSM scaled cross sections for all the ions. %$Z=+1$ to $+8$. , giving similar energy for the maximum of the cross sections. %, as displayed in  figure \ref{fig:zreduced}. 
The validity of the theoretical scaling with the ion charge is very clear in figure \ref{fig:zreduced}, where the CDW-SSM curves lays one over the other.  Out theoretical results are valid for impact energies around and above the maximum of the cross sections, which means above 50 keV for impact of H$^+$  to 250 keV/amu for impact of O$^{+8}$.

The scaling was tested with the experimental data available %for these eight molecules 
by impact of different charged ions \cite{itoh2013,iriki2011,wolff2014,wang2016,tribedi2019,agnihotri2012,agnihotri2013,Luna2007,Rudd86,Rudd85,Luna_Li_water,DalCappello2009,Tribedi_O_water},
and also by electron impact at sufficiently high velocity \cite{rahman2016,bug2017,wolf2019,fuss2009}. As can be noted, most of the data in figure \ref{fig:zreduced} verifies the scaling quite well. The data of uracil by swift C, O and F ions 
in \cite{agnihotri2012,agnihotri2013} are too low as compared with our CDW-SSM results, but also as compared with Itoh \textit{et al} data \cite{itoh2013}, and with the CTMC calculations by Sarkadi \cite{sarkadi2016}.
%as compared to proton impact measurements by Itoh \textit{et al} \cite{itoh2013} multiplied by $Z^2$. Independent CTMC calculations by Sarkadi \cite{sarkadi2016} also mention this. 
The data for Li$^{+3}$ in water in \cite{Luna_Li_water} also separate from the present theoretical curves for $E<600$ keV/amu.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Universal Scaling rule}
%\label{sec:U-scaling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The good results obtained in the scaling with the ion charge challenged us to look for a more general scaling rule that could predict values for ionization cross sections of any ion in any molecule.  % that scales ions and molecules  
%Based on he good results shown in figure \ref{fig:zreduced}  and the scaling with the number of active electrons $n_e$ in each atom in  \cite{MendezJPB20}, we went further for an universal scaling that involves the different molecules biomolecules studied here, and 
To that end, we resorted to the %scaling with the 
number of active electrons in each atom $n_e$ proposed in \cite{MendezJPB20}, and combined it with the $Z$-scaling displayed in figure \ref{fig:zreduced}. %given by Eq. (\ref{Montenegro}) with $\alpha=1.2$.

The CDW ionization cross sections $\sigma^{CDW}$ of atomic H, C, N, O targets scale as $\sigma_e=\sigma^{CDW}/n_e$ with  $n_e$ being $1$ for H, $4$ for C, N, and O. %; and $4.5$ for P and S. 
The SSM leads to the molecular numbers of active electrons included in table \ref{nn}.

\begin{table}[H]
\begin{center}
\begin{tabular}{|ll|ll|ll|}
\hline
 Molecule & $n_e$ &Molecule          & $n_e$ & Molecule             & $n_e$\\
\hline
% --------------------------------------------------------------------------------
% H$_2$    & 2  & CH$_5$N             \ \ \ \ & 13    & C$_4$H$_5$N$_3$O     \ \ \ \ & 37   \\ %H2 -pyridine  -cytosine
% H$_2$O   & 6  & C$_4$H$_8$O         \ \ \ \ & 28    & C$_5$H$_6$N$_2$O$_2$ \ \ \ \ & 42   \\ %H2O-THF       -thymine
% O$_2$    & 8  & C$_4$H$_4$N$_2$    \ \ \ \  & 28    & C$_5$H$_5$N$_5$      \ \ \ \ & 45   \\ %O2- pyrimidine-adenine
% CH$_4$   & 8  & C$_6$H$_6$          \ \ \ \ & 30    & C$_5$H$_5$N$_5$O     \ \ \ \ & 49   \\ %methane-benzene-guanine
% CO$_2$   & 12 & C$_4$H$_4$N$_2$O$_2$\ \ \ \ & 36    & C$_5$H$_{10}$O$_5$P \ \ \ \ & 54.5 
% \\ %CO2- uracil-backbone
%--------------------------------------------------------------------------------
% Nueva tabla
% -------------------------------------------------------------------------------
 H$_2$O   & 6   &  CO$_2$  \ \ \ \ & 12  & C$_4$H$_5$N$_3$O     \ \ \ \ & 37   \\ %H2 -CO2  -cytosine
 N$_2$    & 8  & C$_4$H$_8$O         \ \ \ \ & 28    & C$_5$H$_6$N$_2$O$_2$ \ \ \ \ & 42   \\ %H2O-THF       -thymine
 O$_2$    & 8  & C$_4$H$_4$N$_2$    \ \ \ \  & 28    & C$_5$H$_5$N$_5$      \ \ \ \ & 45   \\ %O2- pyrimidine-adenine
 CH$_4$   & 8  & C$_4$H$_4$N$_2$O$_2$ \ \ \ \ &  36   & C$_5$H$_5$N$_5$O     \ \ \ \ & 49   \\ %methane-uracil-guanine
% --------------------------------------------------------------------------------
 \hline
\end{tabular}
\caption{Number of active electrons per target at intermediate to high energies obtained from the CDW calculations~\cite{MendezJPB20}.}
\label{nn}
\end{center}
\end{table}

\begin{figure*}[!t]%[h]
\centering
\includegraphics[width=0.7\textwidth]{zmol_alpha.eps}
\caption{(Color online) Universal scaling with the ions charge $Z$ and the number of active electrons in the molecule $n_e$ given by Eq. (\ref{u-scaling}). Curves: present CDW-SSM theoretical results. 
Symbols: experimental impact of H$^+$ on 
\mbox{\Large$\circ$} adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\triangledown$ pyrimidine~\cite{wolff2014} and 
$\meddiamond$ THF~\cite{wang2016};
$\lozenge$ C$^{+6}$ on adenine \cite{tribedi2019};
\mbox{\fontsize{10}{20} $\varhexagon$} H$^+$ \cite{Luna2007},
$\pentagon$ H$^+$ \cite{Rudd86}, 
$\odot$ He$^{+2}$ \cite{Rudd85},  
$\ovee$ C$^{+6}$ \cite{DalCappello2009}, and 
$\boxslash$ O$^{+8}$ \cite{Tribedi_O_water} on water;
H$^{+}$ and He$^{+2}$ impact on 
$\boxast$~N$_2$, 
$\boxbox$~O$_2$,
$\boxbar$~CO, 
$\boxbslash$~CO$_2$, and
$\boxdot$~CH$_4$ \cite{Rudd85,Rudd1983}, 
$\varowedge$~H$^{+}$ on CH$_4$~\cite{Luna2019}; 
and electron impact on $\rhd$~pyrimidine~\cite{bug2017}, and $\lhd$, $\medstar$~\cite{wolf2019,fuss2009} THF.}
\label{fig:zalpha}
\end{figure*}

The universal scaling we propose here is expressed as
 \begin{equation}
     \sigma_U=\frac{\sigma_e}{Z^{\alpha}}=\frac{\sigma/n_e}{Z^{\alpha}}\,,
     \label{u-scaling}
 \end{equation}
as a function of $E/Z^{2-\alpha}$, with $\sigma$ being the ionization cross section for the molecular target, $\alpha=1.2$ and  $n_e$ being the number of active electrons per molecule in table \ref{nn}. %It is important to point out that the scaling is based on CDW results atomic H, C, N, O, and P targets, so it is only  valid for ionization cross sections %by charged ions (at least up to +8) of molecules composed by these atoms.%, unless other values for ne were obtained for other atomic targets. 
In figure \ref{fig:zalpha} we test the universal scaling of Eq. (\ref{u-scaling}) %by showing in the same plot 
for all the theoretical and experimental values displayed in figure \ref{fig:zreduced}. 
As can be noted, the universal scaling works fine, all the curves and data lays in a narrow band valid for any charged ion (scaling with Z) in any molecule (scaling with the number of active electrons). We decided not including in this figure the data  for uracil in \cite{agnihotri2012,agnihotri2013}, and for Li on water in \cite{Luna_Li_water}. The discussion about these experimental values exceeds the present work. %is open for future work.

In principle, the \textit{universal} scaling should be valid for different ion-molecule combinations. 
We proved this by including in figure \ref{fig:zalpha} the measurements by Rudd et al \cite{Rudd85,Rudd1983} for H$^{+}$ and He$^{+2}$ in N$_2$, O$_2$, CH$_4$, CO and CO$_2$, and the recent values by Luna et al \cite{Luna2019} for H$^{+}$ in CH$_4$.

The good agreement shown in figure \ref{fig:zalpha} resumes the main result of this work, and holds the validity of present universal scaling. Although the theoretical CDW-SSM results  %Present scaling rule 
are valid for energies above the maximum of the cross sections, the scaling of the experimental data extends to lower impact energies%is more extensive
, as can be noted in figure \ref{fig:zalpha}.
The importance of scaling rules lay on the predictive capability. New measurements for other ions and molecules are expected to reinforce the present proposal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
In this letter, we present an scaling for the ionization cross sections of highly charged ions in biological targets. The scaling rule states the cross sections divided by $Z^{\alpha}$ as a function of the reduced impact energy $E/Z^{2-\alpha}$, with $\alpha=1.2$. The scaling was obtained by means of the CDW-SSM calculations for five different charged ions in eight targets and tested with the available experimental data. An universal scaling rule  is also proposed, which reduced the cross sections with the number of active electrons of the molecule. The universal scaling proved to be valid for a large number of experimental data.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip
%\section*{References}

\begin{thebibliography}{}

% intro --------------------------------------------------------
\bibitem{PhysMed} T. Liamsuwan and H. Nikjoo, Phys. Med. Biol. \textbf{58}  641–672 (2013).

\bibitem{Mohamad2017}
O. Mohamad, B. J. Sishc, J. Saha, A. Pompos, A. Rahimi, M. D. Story, A. J. Davis, D. N. Kim, Cancers \textbf{9}, 66 (2017).

\bibitem{Solov2009} A. V. Solov'yov, E. Surdutovich, E. Scifoni, I. Mishustin, and W. Greiner, Phys. Rev. E \textbf{79}, 011909 (2009);
% https://link.aps.org/doi/10.1103/PhysRevE.79.011909

\bibitem{Denifl2011}
Denifl S., Märk T.D., Scheier P. 
% The Role of Secondary Electrons in Radiation Damage. 
% In Radiation Damage in Biomolecular Systems. Biological and Medical Physics, Biomedical Engineering. 
Eds: García Gómez-Tejedor G., Fuss M. 
Springer, Dordrecht (2012)

\bibitem{water} N. A. Gafur, M.  Sakakibara, S. Sano, K. A. Sera, 
% \textit{Case Study of Heavy Metal Pollution in Water of Bone River by Artisanal Small-Scale Gold Mine Activities in Eastern Part of Gorontalo, Indonesia}, 
Water \textbf{10}, 1507 (2018); doi:10.3390/w10111507.

\bibitem{ferrazdias} D. Benedetti, E. Nunes, M. Sarmento, C. Porto, C. E. Iochims dos Santos, J. Ferraz Dias, J. da Silva,
% \textit{Genetic damage in soybean workers exposed to pesticides: Evaluation with the comet and buccal micronucleus cytome assays},
Mutation Research/Genetic Toxicology and Environmental Mutagenesis,
Volume 752, 28-33 (2013);
% https://doi.org/10.1016/j.mrgentox.2013.01.001.

\bibitem{vera_prl2013} P. de Vera, R. Garcia-Molina,I. Abril, and A. V. Solov’yov, Phys. Rev. Lett.  110, 148104 (2013).

% our ---------------------------------------------------------
\bibitem{MendezJPB20}A. M. P. Mendez, C. C. Montanari, and J. E. Miraglia, J. Phys. B: At. Mol. Opt. Phys.  \textbf{53}, 055201 (2020).

% others -----------------------------------------------
\bibitem{Quinto20} M.A. Quinto, J.M. Monti, C.A. Tachino, P.F. Weck, O.A. Fojón, C. Champion, R.D. Rivarola, 
% \textit{The physics of irradiation of biological matter by ion beams}, 
Rad. Phys. Chem. \textbf{167}, 108337 (2020);
% https://doi.org/10.1016/j.radphyschem.2019.05.027.

\bibitem{ludde2019}
H. J. L\"udde,  M. Horbatsch and T. Kirchner, J. Phys. B: At. Mol. Opt. Phys.  \textbf{52}, 195203 (2019).

\bibitem{ludde2018}
H. J. L\"udde, M. Horbatsch and T. Kirchner, Eur. Phys. J. B \textbf{91}, 99 (2018).

\bibitem{ludde2016}
H. J. L\"udde, A. Achenbach, T. Kalkbrenner, H.-C. Jankowiak and T. Kirchner,
Eur. Phys. J. D \textbf{70}, 82 (2016).

\bibitem{Champion2012} 
C Champion, M E Galassi, O Foj\'{o}n, H Lekadir, J Hanssen, R D Rivarola, P F Weck, A N Agnihotri, S Nandi, and L C Tribedi. 
J. Phys.: Conf. Ser. \textbf{373}, 012004 (2012).

% Z-scaling------------------------------------------------------
\bibitem{janev1980}
R. K. Janev and L. P. Presnyakov 
J. Phys. B: At. Mol. Opt. Phys.  \textbf{13}, 4233 (1980).

\bibitem{dubois13}
R. D. DuBois, E. C. Montenegro and G. M. Sigaud,
AIP Conference Proceeding \textbf{1525}, 679 (2013).

\bibitem{montenegro_pra13} E. C. Montenegro, G. M. Sigaud,  and R. D. DuBois, Phys. Rev. A \textbf{87} 012706 (2013).

% Experimental data ----------------------------------------------

\bibitem{itoh2013} 
A. Itoh, Y. Iriki, M. Imai, C. Champion, and R. D. Rivarola, 
Phys. Rev. A \textbf{88}, 052711 (2013).

\bibitem{iriki2011}
Y. Iriki, Y. Kikuchi, M. Imai, and A. Itoh
Phys. Rev. A \textbf{84}, 052719 (2011).

%\bibitem{Tabet2010} J. Tabet, S. Eden, S. Feil, H. Abdoul-Carime, B. Farizon, M. Farizon, S. Ouaskit, and T. D. Mark, Phys. Rev. A \textbf{82}, 022703 (2010).

\bibitem{wolff2014}
W. Wolff, H. Luna, L. Sigaud, A. C. Tavares, and E. C. Montenegro
J. Chem. Phys. \textbf{140}, 064309 (2014).

\bibitem{wang2016}
M. Wang, B. Rudek, D. Bennett, P. de Vera, M. Bug, T. Buhr, W. Y. Baek, G. Hilgers, H. Rabus, 
Phys. Rev. A \textbf{93}, 052711 (2016).\bibitem{tribedi2019} S. Bhattacharjee, C. Bagdia, M. R. Chowdhury, A. Mandal, J. M. Monti, R. D. Rivarola, and L. C. Tribedi, 
Phys. Rev. A \textbf{100}, 012703(2019).

\bibitem{agnihotri2012}
A. N. Agnihotri, S. Kasthurirangan, S. Nandi, A.
Kumar, M. E. Galassi, R. D. Rivarola, O. Foj\'{o}n, C. Champion, J. Hanssen, H. Lekadir, P. F. Weck, and L. C. Tribedi. 
Phys. Rev. A \textbf{85}, 032711 (2012).

\bibitem{agnihotri2013}
A N Agnihotri, S Kasthurirangan, S Nandi, A Kumar, C Champion,, H Lekadir, J Hanssen, P FWeck, M E Galassi, R D Rivarola, O Fojon and L C Tribedi, 
J. Phys. B: At. Mol. Opt. Phys.  \textbf{46}, 185201 (2013).

% H+ in water-------------------------------------
\bibitem{Luna2007}
H. Luna, A. L. F. de Barros, J. A. Wyer, S. W. J. Scully, J. Lecointre, P. M. Y. Garcia, G. M. Sigaud, A. C. F. Santos,
V. Senthil, M. B. Shah, C. J. Latimer, and E. C. Montenegro,
Phys. Rev. A \textbf{75}, 042711 (2007).

%\bibitem{Champion2007}C. Champion, O. Boudrioua, C. Dal Cappello, Y. Sato, D. Ohsawa, Phys. Rev. A 75 (2007) 032724.

\bibitem{Rudd86} M. A. Bolorizadeh and M. E. Rudd, Phys. Rev. A \textbf{33}, 888 (1986).

% He$^{+2}$ in water---------------------
%\bibitem{DalCappello2007} O. Boudrioua, C. Champion, C. Dal Cappello, Y.V. Popov, Phys. Rev. A 75 (2007) 012717.

\bibitem{Rudd85} M. E. Rudd, T. V. Goffe, and A. Itoh,  Phys. Rev. A \textbf{32}, 2128 (1985).

% Li$^{+3}$ in water---------------------------
\bibitem{Luna_Li_water} H. Luna, W. Wolff, and E. C. Montenegro, Andre C. Tavares, H. J. Ludde, G. Schenk, M. Horbatsch, and T. Kirchner, Phys. Rev. A \textbf{93}, 052705 (2016).

% C$^{+6}$ in water--------------------------
\bibitem{DalCappello2009}
C. Dal Cappello, C. Champion, O. Boudrioua, H. Lekadir, Y. Sato , D. Ohsawa, 
Nuclear Instruments and Methods in Physics Research B 267 (2009) 781–790.

%O$^{+8}$ in water -------------------------------
\bibitem{Tribedi_O_water} S. Bhattacharjee, S. Biswas, C. Bagdia, M. Roychowdhury, S. Nandi, D. Misra, J. M. Monti, C. A. Tachino, R. D. Rivarola, C. Champion and L. C. Tribedi, J. Phys. B: At. Mol. Opt. Phys. \textbf{49,}  065202 (2016).

% electron impact --------------------------
\bibitem{rahman2016}
M. A. Rahman and E. Krishnakumar,
Electron ionization of DNA bases,
J. Chem. Phys. \textbf{144}, 161102 (2016).

\bibitem{bug2017}
M. U. Bug, W. Y. Baek, H. Rabus, C. Villagrasa, S. Meylan, A. B. Rosenfeld,
Rad. Phys. Chem. \textbf{130}, 459--479 (2017).

\bibitem{wolf2019}
W. Wolff, B. Rudek, L. A. da Silva, G. Hilgers, E. C. Montenegro, 
M. G. P. Homem,
J. Chem. Phys. \textbf{151}, 064304 (2019).

\bibitem{fuss2009}
M. Fuss, A. Muñoz, J. C. Oller, F. Blanco, D. Almeida, P. Limão-Vieira, 
T. P. D. Do, M. J. Brunger, G. Garc\'{i}a,
Phys. Rev. A \textbf{80}, 052709 (2009).

\bibitem{sarkadi2016} L Sarkadi, J. Phys. B: At. Mol. Opt. Phys. \textbf{49} 185203 (2016)

% H+ in CH4 and CO2-----------------------------------------
\bibitem{Rudd1983} M. E. Rudd, R. D. DuBois, L. H. Toburen, and C. A. Ratcliffe, T. V. Goffe, Phys. Rev. A \textbf{28}, 3244 (1983).

\bibitem{Luna2019} H. Luna, W. Wolff, and E. C. Montenegro, L. Sigaud, Phys. Rev. A \textbf{99}, 012709 (2019).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%, %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{thebibliography}

\end{document}