### 2020-06-02 scalings/justification.tex

\documentclass[a4paper,10pt]{letter}
\usepackage[utf8]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
\pagestyle{headings}
\date{}% It is always \today, today,

\begin{letter}{}

\opening{}

\begin{center}
\it Justification for Accelerated Publication in Environmental Research Letters
\end{center}

We are submitting electronically a paper entitled:
``Universal scaling for the ionization of biological molecules by 
highly charged ions'' by A.M.P. Mendez, C.C. Montanari and J.E. Miraglia.
The present manuscript is a follow up of previous work [Mendez {\it et al.}
J. Phys B (2020)]. In this letter, we present a scaling for the 
ionization cross sections of highly charged ions in biological targets.
The scaling 
was obtained by means of the CDW--SSM calculations for forty ions--molecule
systems and tested with the available experimental 
data. 
A universal scaling rule is also proposed, which reduced the 
cross sections with the number of active electrons of the molecule. 
The universal scaling proved to be valid for a large number of 
experimental data.

The importance of our scaling rule lays in its predictive capability, 
which is useful in experimental measurements and multipurpose codes.
We believe that our model will appeal to a wide audience, particularly 
for those interested in biological and complex molecules.

\signature{A.M.P. Mendez}

%enclosure listing
%\encl{}

\end{letter}
\end{document}